Climatic warming strengthens a positive feedback between alpine shrubs and fire (Supplementary Material)

### 2016-03-15 manuscript/supplementary.Rnw
Added updated details about DEM

\section*{Seedling survey}
We measured seedling density and maximum heights in open heathland across the Bogong High Plains. The primary aim of this study was to evaluate how variables such as altitude, Topographic Wetness Index (a measure of plant available water), adult density, fire, and fire severity influence the density of \textit{Asterolasia} and \textit{Grevillea} seedlings in open heathland. A secondary aim was to verify growth rates observed in the OTC experiment by assessing maximum seedling heights of \textit{Asterolasia} and \textit{Grevillea} along altitude, soil moisture and fire severity gradients in areas burnt by the 2003 wildfires.

The study used 40 open heathland sites\cite{Camac:2013eb} established immediately after the 2003 fires\cite{Williams:2006tl}. All sites have been free of grazing by cattle since 2003. Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years\cite{Camac:2013eb}. This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires. These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitude from 1660 to 1860 m a.s.l. Sites were at least 0.25 ha and more than 500 m apart. In the summer of 2011-12, 9-years post-fire, seedling density in each site was estimated using forty 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects. Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines. This effectively created a grid of 40 plots (10 plots~\texttimes~4 transects) per site, subsampling an area of 2000 m$^2$.

Different variables were recorded at the three sampling levels (site, transect and plot). For each site, altitude and TWI were derived from a 30 m resolution digital elevation model of the Bogong High Plains. TWI, which approximates the relative plant-available water based on a site\textquotesingle s position in the landscape\cite{Moore:1993fk}, was derived for each site using $\textrm{TWI}= \textrm{ln}(\frac{\alpha}{\textrm{tan}\theta})$, where, $\alpha$ is the upslope catchment area per unit width orthogonal to the flow direction and $\theta$ is the slope (in radians). Across the 40 sites, TWI varied twofold (5 to 10 units; higher units = wetter). Altitude ranged by \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m (i.e. from \Sexpr{round(min(complete_field_survey_data$altitude_m))} to \Sexpr{round(max(complete_field_survey_data$altitude_m))} m a.s.l), or by approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation\cite{Brown:1989vx}. This altitudinal temperature range is comparable to that observed in the OTC experiment.

@@

\section*{Data analysis}
In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package rstan \Sexpr{packageVersion("rstan")}\cite{stan2015xx}. For each model, three independent chains were run. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}. In all cases modelled parameters converged within 1000 iterations. After discarding the first 1000 iterations as ‘warm up’, a further 1000 samples were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. To ensure all covariate effects were identifiable we removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

@@

\begin{equation}\label{eq:mort_bg} 
\begin{aligned}
\lambda_i &=  \alpha_{j[i]} + \beta_{j[i]} \times OTC_{k[i]} + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\alpha_j &\sim \mathcal{N}(\mu_{spp}^{CTL}, \sigma_{spp}^{CTL}) \\
\beta_j &\sim \mathcal{N}(\mu_{spp}^{OTC}, \sigma_{spp}^{OTC}) \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{obs}).
\end{aligned}
\end{equation}

@@

\begin{equation}\label{eq:mort_tussock} 
\begin{aligned}
\lambda_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{obs}).
\end{aligned}
\end{equation}

@@

Here the standard deviation parameters in both mortality models (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{obs}$) were each estimated from positive half Cauchy priors with standard deviation of 2.5. The intercept term $\alpha$ for inter-tussock seedling model and the mean hyper-parameter $\mu_{spp}^{CTL}$ in the multispecies models were estimated from cauchy priors with standard deviations of 2.5. We also used these Cauchy priors for inter-tussock model coefficients (i.e. $\beta_1$, $\beta_2$ and $\beta_3$) and the multi-species hyperparameter for the mean OTC effect $\mu_{spp}^{OTC}$. These cauchy priors are the recommended default choice for binomial regression models\cite{Gelman:2008ua}.

@@

Half Cauchy priors with a scale of 25 were used for both the overdispersion parameter, $\phi$ and the plot level standard deviation $\sigma_{plot}$. Coefficients $\beta_1$ and $\beta_2$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior knowledge that plots were unlikely to have more than 150 recruits/m$^2$.

@@

Half Cauchy priors with a scale of 25 were used for all standard deviation parameters. Hyperparameter $\mu_{site}$ and coefficients $\beta_1$ to $\beta_5$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior belief that plots were unlikely to have more than 150 seedlings/m$^2$.

@@

Here, we estimated $\mu_{site}$ and coefficients ($\beta_1$ to $\beta_3$) using weakly informative normal priors with a mean of 0 and a standard deviation of 2.5. These priors encapsulated our prior knowledge that seedling maximum heights would be less than 150 cm. All standard deviation parameters were modelled using positive half Cauchy priors with a scale of 25.
\pagebreak
\section*{Supplementary figures}
\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS1.pdf}
\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI).} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude and (\textbf{e}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs1}
\end{minipage}
\end{suppfigure}
\pagebreak

>>>>

\section*{Seedling survey}
We measured seedling density and maximum heights in open heathland across the Bogong High Plains. The primary aim of this study was to evaluate how variables such as altitude, Topographic Wetness Index (TWI; a measure of plant available water), adult density, fire, and fire severity influence the density of \textit{Asterolasia} and \textit{Grevillea} seedlings in open heathland. A secondary aim was to verify growth rates observed in the OTC experiment by assessing maximum seedling heights of \textit{Asterolasia} and \textit{Grevillea} along altitude, soil moisture and fire severity gradients in areas burnt by the 2003 wildfires.

The study used 40 open heathland sites\cite{Camac:2013eb} established immediately after the 2003 fires\cite{Williams:2006tl}. All sites have been free of grazing by cattle since 2003. Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years\cite{Camac:2013eb}. This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires. These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitude from \Sexpr{round(min(complete_field_survey_data$altitude_m))} to \Sexpr{round(max(complete_field_survey_data$altitude_m))} m a.s.l. Sites were at least 0.25 ha and more than 500 m apart. In the summer of 2011-12, 9-years post-fire, seedling density in each site was estimated using forty 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects. Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines. This effectively created a grid of 40 plots (10 plots~\texttimes~4 transects) per site, subsampling an area of 2000 m$^2$.

Different variables were recorded at the three sampling levels (site, transect and plot). For each site, altitude and TWI were derived from a 30 m resolution digital elevation model\cite{Gallant:2011er} of the Bogong High Plains. TWI, which approximates the relative plant-available water based on a site\textquotesingle s position in the landscape\cite{Moore:1993fk}, was derived for each site using R package \texttt{dynatopmodel}\cite{Metcalfe:2016ud}. Across the 40 sites, TWI varied twofold (\Sexpr{min(round(complete_field_survey_data$topo_wetness_index))} to \Sexpr{round(max(complete_field_survey_data$topo_wetness_index))} units; higher units = wetter). Altitude ranged by \Sexpr{round(max(complete_field_survey_data$altitude_m)) - round(min(complete_field_survey_data$altitude_m))} m, or by approximately \Sexpr{round(0.77 * (190/100),1)}\textdegree C from the lowest to the highest site (based on a lapse-rate of 0.77\textdegree C/100 m elevation\cite{Brown:1989vx}. This altitudinal temperature range is comparable to that observed in the OTC experiment.

@@

\section*{Data analysis}
In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset\cite{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")}\cite{stan:2015xx}. For each model, three independent chains were run. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic\cite{Gelman:1992ts, Brooks:1998ju}. In all cases modelled parameters converged within 1000 iterations. After discarding the first 1000 iterations as ‘warm up’, a further 1000 samples were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. To ensure all covariate effects were identifiable we removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

@@

\begin{equation}\label{eq:mort_bg} 
\begin{aligned}
\lambda_i &=  \alpha_{j[i]} + \beta_{j[i]} \times OTC_{k[i]} + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\alpha_j &\sim \mathcal{N}(\mu_{spp}^{CTL}, \sigma_{spp}^{CTL}) \\
\beta_j &\sim \mathcal{N}(\mu_{spp}^{OTC}, \sigma_{spp}^{OTC}) \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{ind}).
\end{aligned}
\end{equation}

@@

\begin{equation}\label{eq:mort_tussock} 
\begin{aligned}
\lambda_{i,t} &=  \alpha + \beta_1 \times OTC_{k[i]} + \beta_2 \times Gap_{i,t} + \beta_3 \times (OTC_{k[i]} \times Gap_{i,t}) + \epsilon_{plot_{k[i]}} + \epsilon_i \\
\epsilon_{plot_k} &\sim \mathcal{N}(0, \sigma_{plot}) \\
\epsilon_i &\sim \mathcal{N}(0, \sigma_{ind}).
\end{aligned}
\end{equation}

@@

Here the standard deviation parameters in both mortality models (i.e. $\sigma_{spp}^{CTL}$ , $\sigma_{spp}^{OTC}$, $\sigma_{plot}$ and $\sigma_{ind}$) were each estimated from positive half Cauchy priors with standard deviation of 2.5. The intercept term $\alpha$ for inter-tussock seedling model and the mean hyper-parameter $\mu_{spp}^{CTL}$ in the multispecies models were estimated from cauchy priors with standard deviations of 2.5. We also used these Cauchy priors for inter-tussock model coefficients (i.e. $\beta_1$, $\beta_2$ and $\beta_3$) and the multi-species hyperparameter for the mean OTC effect $\mu_{spp}^{OTC}$. These cauchy priors are the recommended default choice for binomial regression models\cite{Gelman:2008ua}.

@@

Half Cauchy priors with a scale of 25 were used for both the overdispersion parameter, $\phi$ and the plot level standard deviation $\sigma_{plot}$. Coefficients $\beta_1$ and $\beta_2$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior knowledge that plots were unlikely to have more than 100 recruits/m$^2$.

@@

Half Cauchy priors with a scale of 25 were used for all standard deviation parameters. Hyperparameter $\mu_{site}$ and coefficients $\beta_1$ to $\beta_5$ were estimated using normal priors centered on zero with a standard deviation of 2.5, encapsulating our prior belief that plots were unlikely to have more than 100 seedlings/m$^2$.

@@

Here, we estimated $\mu_{site}$ and coefficients ($\beta_1$ to $\beta_3$) using weakly informative normal priors with a mean of 0 and a standard deviation of 2.5. These priors encapsulated our prior knowledge that seedling maximum heights would be less than 100 cm. All standard deviation parameters were modelled using positive half Cauchy priors with a scale of 25.
\pagebreak
\section*{Supplementary figures}
\begin{suppfigure}
\begin{minipage}{\textwidth}
\includegraphics{figures/figS1.pdf}
\caption{\textbf{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI).} (\textbf{a}) Centered and standardized model coefficients (on the log scale); and effects of (\textbf{b}) burning, (\textbf{c}) fire severity, (\textbf{d}) altitude and (\textbf{e}) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs1}
\end{minipage}
\end{suppfigure}
\pagebreak
